Deep Neural Architectures for Discourse Segmentation in E-Mail Based Behavioral Interventions

### 2018-11-26 paper-amia-summit-2019/manuscript.tex
finalize camera ready version of the submitted manuscript

In this study, we focus on the analysis of email-delivered MI, or e-Coaching, to promote healthy eating among young adults. The e-Coaching dataset is composed of email correspondence between a MI counselor and the young adult patient. Unlike transcribed in-person exchanges, email correspondence is not clearly segmented into codable speech acts (i.e., utterances). Thus, the unstructured nature of e-Coaching exchanges poses a unique set of analytic challenges. Segmentation of e-Coaching exchanges into textual fragments that correspond to distinct e-Coach and patient communication behaviors is a significant barrier to qualitative analysis of this type of clinical conversation. Automating this task is a unique and challenging problem due to the following reasons:

@@

\section*{Materials and methods}
\subsection*{\textit{Dataset}}
The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include 330 and 281 emails by e-Coaches and patients, respectively. Various statistics of the experimental dataset are provided in Table~\ref{tab:datastat}. Each e-Coaching session represents a MI intervention delivered via email. Emails were segmented into 3,138 text fragments and annotated with the Minority Youth-Sequential Coding of Process Exchanges (MY-SCOPE),~\cite{carcone2013provider} a qualitative coding scheme to characterize patient-provider communication during MI sessions with 115 distinct behavior codes. Email segmentation can be considered as sequence tagging, which can be framed as a binary classification problem, in which each word or punctuation mark is annotated with one of the two class labels (``new segment'' or ``same segment'') to indicate whether it is a beginning of a new MI behavior segment or not. In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances. In this study, we experimented with traditional machine learning methods, such as Conditional Random Fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bi-directional Recurrent Neural Network (BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} In the case of MLP, training and testing samples were created based on a sliding window of $2n$ words or punctuation marks over each position (which could be a word or a punctuation mark) in a given input sequence, such that each sample consists of the $n$ words or punctuation marks after the current position and $n$ words or punctuation marks prior to the current position, including the position itself. In the case of CRF, BRNN and CRNN models, an e-Coaching email was taken as an input sequence, POS tags and embeddings of each word or punctuation mark were used as input and binary labels corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output. In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1 (an example of an input with the assigned labels is provided in the $2^{nd}$ row of Table~\ref{tab:datastat}).

\begin{table}[ht]
\centering
\caption{\textbf{Summary of statistics of the experimental dataset and example of segmented sequence.}}
\label{tab:datastat}
 \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Sessions}} & \multirow{2}{*}{\textbf{Instances}} & \multicolumn{2}{|c|}{\textbf{Class labels}} & \multicolumn{2}{|c|}{\textbf{Tokens}} & \multicolumn{2}{|c|}{\textbf{Emails}} & \multicolumn{2}{|c|}{\textbf{Annotation}} \\\cline{3-10}
   &  & \textbf{new}  & \textbf{same} & \textbf{words} & \textbf{punctuations}  & \textbf{patient} & \textbf{provider} & \textbf{method}  & \textbf{codes} \\ \hline    
 49 & 102,917 & 3,138 & 99,779 & 95,777 & 7,140 & 281 & 330 & MYSCOPE & 115 \\ \hline
 \multicolumn{10}{|c|}{segmentation example: Hi[0] XXX[0] ,[0] its[0] good[0] to[0] hear[0] from[0] you[0] .[1] it[0] sounds[0] like[0]...} \\ \hline
  \end{tabular}
\end{table}

@@

\textbf{Convolutional Recurrent Neural Network (CRNN)}: CRNN \cite{treviso2017sentence} shown in Figure~\ref{fig:crnn} is a deep neural architecture, which combines convolutional and recurrent layers. Our implementation of CRNN consists of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output. E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input and embedding layers to produce a $m \times n_e$ matrix after fetching the embeddings for words and punctuations in the input sequence. This matrix is a distributed representation of an input email exchange, which contains rich morpho-syntactic information that can be utilized for its segmentation. When POS tags are utilized along with word embeddings, they are represented with a 10-dimensional vector, which is concatenated with 300-dimensional word embeddings to obtain new embedding vectors $n_e = [n_w;n_p]$ of size 310. The primary purpose of a convolution layer is to extract new features for each word or punctuation mark based on the neighboring words or punctuation marks. A one-dimensional (1D) convolution operation is utilized in this layer in our implementation of BRNN. In 1D convolution, one filter is responsible for extraction of one feature. After applying $n_f$ different filters with zero-padding on both sides of the input text, $n_f$ features are produced by the convolution layer for each word. A max pooling over time operation is then applied to find the most significant features in a textual fragment. The bi-directional recurrent layer receives new features extracted by the convolution layer. Unidirectional RNNs are typically utilized to capture long-range dependencies in a sequence of observations. Bi-directional RNNs, on the other hand, are capable of capturing both past and future contexts through forward and backward traversals of a sequence. The purpose of the fully connected layer in CRNN is to use the output of the bidirectional RNN layer for classifying each word or punctuation into ``new segment'' or ``same segment'' classes. Since a fully connected layer has a larger number of parameters, they are more likely to excessively co-adapt to other parameters in the network and result in over-fitting. To prevent this, we utilized dropout by randomly ignoring 50\% of the connections in the fully connected layer of CRNN. Finally, logistic sigmoid outputs the probability of classifying or labeling each word or punctuation mark with ``same segment'' class. We experimentally determined the optimal parameters using 5-fold cross-validation and found out that the best performance is achieved when filter length in the convolution layer is 7, number of filters is 100, max pooling size is 3, ReLU is used as an activation function in the convolution layer, hyperbolic tangent is used as an activation function in the bi-directional RNN layer and the number of dimensions in the hidden state of RNNs is 200. Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{source code of all methods is available at \url{https://github.com/teanalab/eCoaching-Text-Segmentation}}

@@

Figure~\ref{fig:embedding-dimension-mlp} (left) illustrates the performance of CRNN on the task of e-Coaching email segmentation by varying the number of dimensions in pre-trained and corpus-based GloVe \footnote{https://nlp.stanford.edu/projects/glove/} and word2vec embeddings. We observed that the best performance is achieved with pre-trained 300-dimensional word2vec word vectors, when three types of features are used together. Therefore, we report the results for other deep learning models used in this study when 300-dimensional word2vec embedding vectors are utilized. The input layer of MLP consists of a sum of embeddings of $n$ words or punctuation marks before and the sum of embeddings of $n$ words or punctuation marks after the word or punctuation mark, which is the center of a sliding window of $2n$ words or punctuation marks. Figure~\ref{fig:embedding-dimension-mlp} (right) demonstrates the performance of MLP on e-Coaching email segmentation by varying the size of the sliding window. It can be observed that the best performance of MLP is achieved when the size of the sliding window is 4 (or $n=2$). Therefore, MLP results in the remaining experiments are reported when $n$ is set to 2.

\begin{table}[ht]
\centering
\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN on ``new segment'' detection as well as the weighted average over ``new segment'' and ``same segment'' classes when only lexical features are used. The highest value for each performance metric is highlighted in boldface.}}
\label{tab:result_base}
  \begin{tabular}{|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} \\\cline{2-7}
   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
 CRF & 0.782 & 0.691 & 0.733 & 0.983 & 0.984 & 0.984 \\ \hline
 MLP & \textbf{0.836} & 0.593 & 0.694 & 0.982 & 0.983 & 0.982 \\ \hline
 BRNN & 0.606 & 0.680 & 0.641 & 0.977 & 0.976 & 0.976 \\ \hline
 CRNN & 0.775 & \textbf{0.797} & \textbf{0.785} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986} \\ \hline
  \end{tabular}
\end{table}

@@

\begin{table}[ht]
\centering
\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN on ``new segment'' detection as well as the weighted average over ``new segment'' and ``same segment'' classes when all types of features are used together. The highest value for each performance metric is highlighted in boldface.}}
\label{tab:result_weighted_avg}
 \begin{tabular}{|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} \\\cline{2-7}
   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} \\ \hline    
 CRF & 0.813 & 0.772 & 0.792 & 0.988 & 0.988 & 0.988 \\ \hline
 MLP & \textbf{0.817} & 0.710 & 0.760 & 0.986 & 0.987 & 0.986 \\ \hline
 BRNN & 0.683 & 0.820 & 0.745 & 0.985 & 0.983 & 0.984 \\ \hline
 CRNN & 0.789 & \textbf{0.864} & \textbf{0.825} & \textbf{0.990} & \textbf{0.989} & \textbf{0.989} \\ \hline
  \end{tabular}
\end{table}

Table~\ref{tab:result_weighted_avg} summarizes the results of all models on the task of segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features. Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the best performance among all methods achieving 0.864 recall with 0.825 F1-measure for ``new segment'' detection and 0.990 precision with 0.989 recall and F1-measure overall. BRNN and CRF demonstrated the lowest and second highest performance on the task of email segmentation among all methods, respectively. We observed that classification performance significantly improved for ``new segment'' detection when lexical features are used in combination with punctuation and POS features. Specifically, precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, on new segment detection when all types of features are utilized together. Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, as a weighted average over ``new segment'' and ``same segment'' classes when lexical features are used in combination with punctuation and POS features.\\

@@

\section*{Discussion}
This study is the first effort to design and evaluate machine learning methods for automated segmentation of e-Coaching sessions. Experimental results indicate that CRNN is the best model among all machine learning methods considered for this study. CRNN achieved 0.989 F1-measure overall and 0.825 F1-measure for detecting ``new segment''. Robust performance of CRNN provides an evidence that deep learning models are capable of detecting the boundaries of patient and provider behaviors in email delivered behavioral interventions. Our experiments also highlight the importance of punctuation and POS features along with word embeddings for all machine learning methods employed this study. Although the domain of this study was intentionally focused, we believe that the proposed methods are not limited to e-Coaching sessions and our conclusions can be generalized to other domains, which require discourse segmentation.

Punctuation marks and POS features resulted in significant improvement in the performance of traditional machine learning and deep learning methods. Punctuation features had stronger individual impact on model performance than POS features. In all cases, CRF and MLP performed better, when word embeddings were used in conjunction with punctuations and POS features. Considering punctuations improved the performance of BRNN and CRNN measured by precision, recall and F1-measure, while POS features lowered their AUPR. %We believe that BRNN and CRNN performed poorly with POS features, since POS taggers already utilize previous and following word as features. Since neighboring words are also considered by bi-directional RNN, such redundancy did not contribute to better performance. %We observed that MLP achieved the highest precision which may be related to the fact that MLP poorly learned ``new segment'' and misclassified new segment words to same segments in 30\%-40\% of the time.

@@

Segmentation is the first step of qualitative analysis of unstructured clinical communications, such as e-Coaching. Although several studies have focused on segmentation problem in biomedical context, they are limited to segmenting clinical text in EHR into sections and sentences. No previous studies considered the task of automated segmentation of clinical communications into groups of MI behaviors in the context of unstructured MI sessions. By comparing the performance of machine learning methods for the task of segmentation of e-Coaching emails, we found out that convolutional recurrent neural networks demonstrate the best performance in terms of most performance metrics. Manual segmentation of e-Coaching sessions is a very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop effective behavioral interventions. Our proposed methods can help to identify textual segments corresponding to MI behaviors in unstructured clinical dialog, which can then be annotated with MI behavior annotation methods in a pipeline setting. Automated segmentation and annotation of e-Coaching emails can significantly decrease the time to identify effective communication strategies in email-based MI. As future work, we are planning to evaluate the methods presented in this work on the datasets from other types of behavioral interventions.

>>>>

In this study, we focus on the analysis of email-delivered MI, or e-Coaching, to promote healthy eating among young adults. The e-Coaching dataset is composed of email correspondence between an MI counselor and the young adult patient. Unlike transcribed in-person exchanges, email correspondence is not clearly segmented into codable speech acts (i.e., utterances). Thus, the unstructured nature of e-Coaching exchanges poses a unique set of analytic challenges. Segmentation of e-Coaching exchanges into textual fragments that correspond to distinct e-Coach and patient communication behaviors is a significant barrier to qualitative analysis of this type of clinical conversation. Automating this task is a unique and challenging problem due to the following reasons:

@@

\section*{Materials and methods}
\subsection*{\textit{Dataset}}
The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include 330 and 281 emails by e-Coaches and patients, respectively. Various statistics of the experimental dataset are provided in Table~\ref{tab:datastat}. Each e-Coaching session represents an MI intervention delivered via email. Emails were segmented into 3,138 text fragments and annotated with the Minority Youth-Sequential Coding of Process Exchanges (MY-SCOPE),~\cite{carcone2013provider} a qualitative coding scheme to characterize patient-provider communication during MI sessions with 115 distinct behavior codes. Email segmentation can be considered as sequence tagging, which can be framed as a binary classification problem, in which each word or punctuation mark is annotated with one of the two class labels (``new segment'' or ``same segment'') to indicate whether it is a beginning of a new MI behavior segment or not. In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances. In this study, we experimented with traditional machine learning methods, such as Conditional Random Fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bi-directional Recurrent Neural Network (BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} In the case of MLP, training and testing samples were created based on a sliding window of $2n$ words or punctuation marks over each position (which could be a word or a punctuation mark) in a given input sequence, such that each sample consists of the $n$ words or punctuation marks after the current position and $n$ words or punctuation marks prior to the current position, including the position itself. In the case of CRF, BRNN and CRNN models, an e-Coaching email was taken as an input sequence, POS tags and embeddings of each word or punctuation mark were used as input and binary labels corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output. In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1 (an example of an input with the assigned labels is provided in the $2^{nd}$ row of Table~\ref{tab:datastat}).

\begin{table}[ht]
\centering
\caption{\textbf{Summary of statistics of the experimental dataset and example of a segmented sequence}}
\label{tab:datastat}
 \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Sessions}} & \multirow{2}{*}{\textbf{Instances}} & \multicolumn{2}{|c|}{\textbf{Class labels}} & \multicolumn{2}{|c|}{\textbf{Tokens}} & \multicolumn{2}{|c|}{\textbf{Emails}} & \multicolumn{2}{|c|}{\textbf{Annotation}} \\\cline{3-10}
   &  & \textbf{new}  & \textbf{same} & \textbf{words} & \textbf{punctuations}  & \textbf{patient} & \textbf{provider} & \textbf{method}  & \textbf{codes} \\ \hline    
 49 & 102,917 & 3,138 & 99,779 & 95,777 & 7,140 & 281 & 330 & MYSCOPE & 115 \\ \hline
 \multicolumn{10}{|c|}{segmentation example: Hi[0] XXX[0] ,[0] its[0] good[0] to[0] hear[0] from[0] you[0] .[1] it[0] sounds[0] like[0]...} \\ \hline
  \end{tabular}
\end{table}

@@

\textbf{Convolutional Recurrent Neural Network (CRNN)}: CRNN \cite{treviso2017sentence} shown in Figure~\ref{fig:crnn} is a deep neural architecture, which combines convolutional and recurrent layers. Our implementation of CRNN consists of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output. E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input and embedding layers to produce a $m \times n_e$ matrix after fetching the embeddings for words and punctuations in the input sequence. This matrix is a distributed representation of an input email exchange, which contains rich morpho-syntactic information that can be utilized for its segmentation. When POS tags are utilized along with word embeddings, they are represented with a 10-dimensional vector, which is concatenated with 300-dimensional word embeddings to obtain new embedding vectors $n_e = [n_w;n_p]$ of size 310. The primary purpose of a convolution layer is to extract new features for each word or punctuation mark based on the neighboring words or punctuation marks. A one-dimensional (1D) convolution operation is utilized in this layer in our implementation of BRNN. In 1D convolution, one filter is responsible for the extraction of one feature. After applying $n_f$ different filters with zero-padding on both sides of the input text, $n_f$ features are produced by the convolution layer for each word. A max pooling over time operation is then applied to find the most significant features in a textual fragment. The bi-directional recurrent layer receives new features extracted from the convolution layer. Unidirectional RNNs are typically utilized to capture long-range dependencies in a sequence of observations. Bi-directional RNNs, on the other hand, are capable of capturing both past and future contexts through forward and backward traversals of a sequence. The purpose of the fully connected layer in CRNN is to use the output of the bidirectional RNN layer for classifying each word or punctuation into ``new segment'' or ``same segment'' classes. Since a fully connected layer has a larger number of parameters, they are more likely to excessively co-adapt to other parameters in the network and result in over-fitting. To prevent this, we utilized dropout by randomly ignoring 50\% of the connections in the fully connected layer of CRNN. Finally, logistic sigmoid outputs the probability of classifying or labeling each word or punctuation mark with ``same segment'' class. We experimentally determined the optimal parameters using 5-fold cross-validation and found out that the best performance is achieved when filter length in the convolution layer is 7, number of filters is 100, max pooling size is 3, ReLU is used as an activation function in the convolution layer, hyperbolic tangent is used as an activation function in the bi-directional RNN layer and the number of dimensions in the hidden state of RNNs is 200. Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{source code of all methods is available at \url{https://github.com/teanalab/eCoaching-Text-Segmentation}}

@@

Figure~\ref{fig:embedding-dimension-mlp} (left) illustrates the performance of CRNN on the task of e-Coaching email segmentation by varying the number of dimensions in pre-trained and corpus-based GloVe \footnote{\url{https://nlp.stanford.edu/projects/glove/}} and word2vec embeddings. We observed that the best performance is achieved with pre-trained 300-dimensional word2vec word vectors, when three types of features are used together. Therefore, we report the results for other deep learning models used in this study when 300-dimensional word2vec embedding vectors are utilized. The input layer of MLP consists of a sum of embeddings of $n$ words or punctuation marks before and the sum of embeddings of $n$ words or punctuation marks after the word or punctuation mark, which is the center of a sliding window of $2n$ words or punctuation marks. Figure~\ref{fig:embedding-dimension-mlp} (right) demonstrates the performance of MLP on e-Coaching email segmentation by varying the size of the sliding window. It can be observed that the best performance of MLP is achieved when the size of the sliding window is 4 (or $n=2$). Therefore, MLP results in the remaining experiments are reported when $n$ is set to 2.

\begin{table}[ht]
\centering
\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN on ``new segment'' detection as well as the weighted average over ``new segment'' and ``same segment'' classes when only lexical features are used. The highest value for each performance metric is highlighted in boldface.}}
\label{tab:result_base}
  \begin{tabular}{|l|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} & \multirow{2}{*}{\textbf{AUPR}} \\\cline{2-7}
   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \\ \hline    
 CRF & 0.782 & 0.691 & 0.733 & 0.983 & 0.984 & 0.984 & 0.780\\ \hline
 MLP & \textbf{0.836} & 0.593 & 0.694 & 0.982 & 0.983 & 0.982 & 0.736 \\ \hline
 BRNN & 0.606 & 0.680 & 0.641 & 0.977 & 0.976 & 0.976 & 0.655\\ \hline
 CRNN & 0.775 & \textbf{0.797} & \textbf{0.785} & \textbf{0.986} & \textbf{0.986} & \textbf{0.986} & \textbf{0.818} \\ \hline
  \end{tabular}
\end{table}

@@

\begin{table}[ht]
\centering
\caption{\textbf{Performance of CRF, MLP, BRNN and CRNN on ``new segment'' detection as well as the weighted average over ``new segment'' and ``same segment'' classes when all types of features are used together. The highest value for each performance metric is highlighted in boldface.}}
\label{tab:result_weighted_avg}
 \begin{tabular}{|l|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Method}} & \multicolumn{3}{|c|}{\textbf{New Segment}} & \multicolumn{3}{|c|}{\textbf{Overall}} & \multirow{2}{*}{\textbf{AUPR}} \\\cline{2-7}
   & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \textbf{Precision}  & \textbf{Recall} & \textbf{F1-Measure} & \\ \hline    
 CRF & 0.813 & 0.772 & 0.792 & 0.988 & 0.988 & 0.988 & \textbf{0.877} \\ \hline
 MLP & \textbf{0.817} & 0.710 & 0.760 & 0.986 & 0.987 & 0.986 & 0.842 \\ \hline
 BRNN & 0.683 & 0.820 & 0.745 & 0.985 & 0.983 & 0.984 & 0.770 \\ \hline
 CRNN & 0.789 & \textbf{0.864} & \textbf{0.825} & \textbf{0.990} & \textbf{0.989} & \textbf{0.989} & 0.867 \\ \hline
  \end{tabular}
\end{table}

Table~\ref{tab:result_weighted_avg} summarizes the results of all models on the task of segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features. Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the best performance among all methods achieving 0.864 recall with 0.825 F1-measure for ``new segment'' detection and 0.990 precision with 0.989 recall and F1-measure overall. BRNN and CRF demonstrated the lowest and second highest performance on the task of email segmentation among all methods, respectively. We observed that classification performance significantly improved for ``new segment'' detection when lexical features are used in combination with punctuation and POS features. Specifically, precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, on new segment detection when all types of features are utilized together. Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, as a weighted average over ``new segment'' and ``same segment'' classes when lexical features are used in combination with punctuation and POS features.

@@

\section*{Discussion}
This study is the first effort to design and evaluate machine learning methods for automated segmentation of e-Coaching sessions. Experimental results indicate that CRNN is the best model among all machine learning methods considered for this study. CRNN achieved 0.989 F1-measure overall and 0.825 F1-measure for detecting ``new segment''. The robust performance of CRNN provides an evidence that deep learning models are capable of detecting the boundaries of patient and provider behaviors in email delivered behavioral interventions. Our experiments also highlight the importance of punctuation and POS features along with word embeddings for all machine learning methods employed this study. Although the domain of this study was intentionally focused, we believe that the proposed methods are not limited to e-Coaching sessions and our conclusions can be generalized to other domains, which require discourse segmentation.

Punctuation marks and POS features resulted in significant improvement in the performance of traditional machine learning and deep learning methods. Punctuation features had a stronger individual impact on model performance than POS features. In all cases, CRF and MLP performed better, when word embeddings were used in conjunction with punctuations and POS features. Considering punctuations improved the performance of BRNN and CRNN measured by precision, recall and F1-measure, while POS features lowered their AUPR. %We believe that BRNN and CRNN performed poorly with POS features, since POS taggers already utilize previous and following word as features. Since neighboring words are also considered by bi-directional RNN, such redundancy did not contribute to better performance. %We observed that MLP achieved the highest precision which may be related to the fact that MLP poorly learned ``new segment'' and misclassified new segment words to same segments in 30\%-40\% of the time.

@@

Segmentation is the first step of qualitative analysis of unstructured clinical communications, such as e-Coaching. Although several studies have focused on the segmentation problem in biomedical context, they are limited to segmenting clinical text in EHR into sections and sentences. No previous studies considered the task of automated segmentation of clinical communications into groups of MI behaviors in the context of unstructured MI sessions. By comparing the performance of machine learning methods for the task of segmentation of e-Coaching emails, we found out that convolutional recurrent neural networks demonstrate the best performance in terms of most performance metrics. Manual segmentation of e-Coaching sessions is a very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop effective behavioral interventions. Our proposed methods can help to identify textual segments corresponding to MI behaviors in unstructured clinical dialog, which can then be annotated with MI behavior annotation methods in a pipeline setting. Automated segmentation and annotation of e-Coaching emails can significantly decrease the time to identify effective communication strategies in email-based MI. As future work, we are planning to evaluate the methods presented in this work on the datasets from other types of behavioral interventions.
